---
title: "Project 7"
author: "Marisa Millenson, Olivia King, Becca Wang, Tammy Jiang, Menghan Hu, Masahiro Nakanishi, Zohaib Siddique"
date: "Due October 26, 2016 at 5:00pm"
output:
  html_document


---

<style type="text/css">
.table {

    width: 80%;
    margin-left:10%; 
    margin-right:10%;
}
</style>
```{r,setup, echo=FALSE, cache=TRUE}
## numbers >= 10^5 will be denoted in scientific notation,
## and rounded to 2 digits
options(scipen = 3, digits = 3)
```




#Project Goals:

The goal of this project is to get better  aquainted with Dplyr and working with data. 

[Project 7 Markdown]()

# New York City Flights 13

- `flights` [336,776 x 16]. Every flight leaving NYC in 2013.
- `weather` [ 8719 x 14]. Hourly Weather Data.
- `planes` [3322 x 9]. Plane data.
- `airports` [1397 x 7]. Airport data. 


```{R}
library(nycflights13)
library(dplyr)
flights
```


# Basic Dyplyr

The deplyr package uses verbs to perform data manipulation.
- `filter()` and `slice()`
- `arrange()`
- `select()` and `rename()`
- `distinct()`
- `mutate()` and `transmute()`
- `sample_n()` and `sample_frac()`

## Filtering Rows

1. Filter all flights on your birthday. 
```{r}
flights %>%
  filter(month==12, day==29)
```

2. Filter all flights in rows 567 - 589
```{r}
flights %>%
  slice(567:589)
```


## Arranging Rows

1. Use  `arrange()` to order the data by `year, month, and day`
```{r,}
flights %>%
  arrange(year,month,day)
```


2. Use `desc()` to arrange flights by arrival delay. 
```{r}
flights %>%
  arrange(desc(arr_delay))
```

## Select

1. Select columns containing date information.

```{r}
flights %>%
  select(year,month,day,hour,minute,time_hour)
```

2. Select columns including only plane information.

```{r}
flights %>%
  select(carrier,flight,tailnum,origin, dest, air_time, distance)
```

3. Rename `tailnum` to `tail_num`

```{r}
flights %>% rename(tail_num = tailnum)
```


## Mutating Rows

1. Create a variable called `gain` which describes the difference between arrival delap and departure delay.

```{r}
flights %>%
  mutate(gain=arr_delay - dep_delay)
```

2. Create a variable called speed which is distance over time in hours. 

```{r}
flights %>%
  mutate(speed=distance/hour)
```

## Summarise

1. Summarise the average delay for all flights.

```{r}
flights %>%
  summarise(mean(dep_delay+arr_delay, na.rm = T))
```


# Overall Questions 

1. Split the complete dataset into individual planes and then summarise each plane by counting the number of flights (count = n()) and computing the average distance and arrival delay .  Then plot this:

```{r}
delay <- flights %>%
  group_by(carrier,tailnum) %>%
  summarise(count = n(), dist = mean(distance, na.rm = T), delay = mean(arr_delay, na.rm = T))
 
library(ggplot2)  
ggplot(delay, aes(dist, delay)) +
  geom_point(aes(size = count), alpha = 1/2) +
  geom_smooth() +
  scale_size_area()
```


2.  Find the number of planes and the number of flights that go to each possible destination. 

```{r}
flights %>%
  group_by(dest) %>%
  summarise(count = n(), planecount = n_distinct(tailnum, carrier))
```


### More Difficult

In the videos we learned about piping our commands. The goal of piping is to help use with the flow of our analysis. The steps we take are:
- Think about it. 
- Describe it. 
- Do it. 


### Consider the function below:



    ```{r}
    result1 <- filter(
      summarise(
        select(
          group_by(flights, year, month, day),
          arr_delay, dep_delay
        ),
        arr = mean(arr_delay, na.rm = TRUE),
        dep = mean(dep_delay, na.rm = TRUE)
      ),
      arr > 30 | dep > 30
    )
    ```



  a. What are some problems with this function? 

This function is hard to read and to follow the logic of-- it also follows some non-standard formatting conventions that I'm not familiar with.
  
  b. How easy is it to follow the logic of this? 
  
    Not very.
    
  c. Rewrite this using piping to make it more understandable. 
```{r}
result2 <- flights %>%
  group_by(year,month,day) %>%
  summarize(arr=mean(arr_delay,na.rm=T), dep=mean(dep_delay,na.rm=T)) %>%
   filter(arr>30| dep>30)
```
  d. Use your function testing skills to verify they return the same results. 
  
```{r}
all.equal(result1, result2)
```

  e.  How would you write the above in terms of SQL?
  
```{r, eval=FALSE}
SELECT year, month, day, avg(arr_delay) AS arr, avg(dep_delay) AS dep FROM FLIGHTS GROUP_BY year, month, day HAVING arr > 30 OR dep > 30
```
  
  
### Consider the function below



```{r}
result3 <- filter(
            summarise(
                 group_by(
                    filter(
                      flights, 
                       !is.na(dep_delay)
                       ),
                       day, hour
                      ),
                       delay=mean(dep_delay),
                      n=n()
                      ),
                       n>10
                       )
                                          
```


  a. What are some problems with this function? 
  Hard to understand what is happening and the order in which things are taking place. 
  
  b. How easy is it to follow the logic of this? 
  Not very. The placement of the brackets make it difficult to comprehend which variables are being used in each function.
  
  c. Rewrite this using piping to make it more understandable.
```{r}
result4 <- flights %>%
  filter(!is.na(dep_delay)) %>%
  group_by(day, hour) %>%
  summarise(delay=mean(dep_delay), n=n()) %>%
  filter(n >10)
```
 
  d. Use your function testing skills to verify they return the same results. 
```{r}
all.equal(result3, result4)
```
  
  e.  How would you write the above in terms of SQL?

```{r, eval=FALSE}
Select day, hour, avg(dep_delay) as delay, count(*) as n
from flights
where dep_delay is not NULL
group by day, hour
having n>10
```

## Consider how to join information from the other datasets in the package with the current information. Is there a way to tell if weather affected the departure  time? Ask yourself other questions about this data and then answer them below. 

The weather dataset includes variables such as wind direction, wind speed, wind gust, precipitation, pressure, and visibility. We can run a factor analysis containing these variables. We can then run a linear regression with departure time as the outcome variable and include the results from the factor analysis and other possible variables that may affect departure time to see if weather affects departure time.

Which airports have the most and least delay?

Most:

```{r}
flights %>%
    group_by(origin) %>%
    summarise(avgdelay=mean(dep_delay, na.rm=T)) %>%
    arrange(desc(avgdelay))
```

Least:

```{r}
flights %>%
    group_by(dest) %>%
    summarise(avgdelay=mean(dep_delay, na.rm=T)) %>%
    arrange(avgdelay)
```

Bring in the weather data by joining the tables. How much does the weather data explain the delay results?

```{r}
joined <- inner_join(flights, weather)
joined %>%
  group_by(origin) %>%
  summarise(avgprecip=mean(precip, na.rm=T), avgwindspeed = mean(wind_speed, na.rm=T), avgvisib=mean(visib, na.rm=T), avgpressure=mean(pressure, na.rm=T), avgwinddir=mean(wind_dir, na.rm=T))
```