---
title: "Project 6: part b"
author: "Liv King, Altan Allawala, Menghan Hu, Becca Wang"
date: "April 14, 2016"
output:
  html_document
---

```{r setup, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(eval=FALSE)

```

<style type="text/css">
.table {

    width: 80%;
    margin-left:10%; 
    margin-right:10%;
}
</style>

####[Lab 7 Markdown]()

## Setup 

For the setup of this, I have created a database called titanic. The goal will be to use this lab to work on the titanic dataset. The code below will let you connect to the database. 

```{R}
library(RMySQL)
con <- dbConnect(MySQL(),
                 user = 'trial',
                 password = 'password',
                 host = '127.0.0.1',
                 dbname='titanic')
```


1. What tables are in this database?


2. Load the titanic data:

```{r}
titanic <- read.table("/home/php2560/titanic.csv", header=TRUE, sep=",")
```

3. Create a table in MySQL from this data. Name it `titanic_yourlastname`:

```{r}
dbWriteTable(con, "titanic_king", titanic)
```

4. Now what tables are in this database?



## Basic usage of commands

We will start to use a list of commands that would be commonly used in MySQL:

These are:

- show tables;
- describe titanic_yourlastname;
- select name, pclass, age, sex from titanic limit 10;
 


We can run any of these commands as follow:

```
query <- "show tables;"
dbGetQuery(con, query)
```

Try these commands and test this out. 


## R and MySQL

5. Run the following SQL command. Before doing so describe what this is asking

Pulling columns of passenger class, whether they survived, and their average age (which is renamed) from the table and then grouping by class and survival.

```{r}
query <- "select pclass, survived, avg(age) as avg_age from titanic_king   
   group by pclass, survived;"
dbGetQuery(con, query)

```


6.  Run the following SQL command. Before doing so describe what this is asking

We are modifying the preexisting table to only display passengers who survived by passenger class.

```{r}
query <- "select pclass, survived, avg(age) as avg_age from titanic_king
   where survived=1
   group by pclass, survived;"
dbGetQuery(con, query)
```


7. Run the following SQL command. Before doing so describe what this is asking

We are selecting the columns name, passenger class, whether or not they survived, and age from titanic_king for those records where names that 1) have the three letters "sen" in the middle; 2) begin with any one character before "sen"; and 3) ends with at least another character. 

```{r}
query <-"select name, pclass, survived, age from titanic_king 
   where name regexp '^.sen.+';"
dbGetQuery(con, query)
```

8. Run the following SQL command. Before doing so describe what this is asking

We are selecting the same columns for those records where the name begins with Jakob and may or may not have more letters after Jakob. 

```{r}
query <- "select name, pclass, survived, age from titanic_king 
   where name regexp 'Jakob.*$';"
dbGetQuery(con, query)
```

9.  Run the following SQL commands. What is different about them? What do they return?
```{r}
query <- "select name, pclass, survived, age from titanic_king 
   where name regexp 'Ivar.*$';"
  
query2 <- "select name, pclass, survived, age from titanic_king 
   where name regexp ',.*Ivar.*$';"

dbGetQuery(con, query)
dbGetQuery(con, query2)
```
The first query returns two names, one of which contains "ivar" as part of a longer name (Givard). The second query returns only one name, which has "Ivar" as a complete word and not within another. In reality, what the second query is doing is specifying that a comma must appear before the "Ivar" string, which means that it will only look within first names.

10. We can also plot data from this:
```{r}
myQuery <- "select pclass, avg(fare) as avg_fare from titanic_king
              group by pclass;"
myData <- dbGetQuery(con, myQuery)
library(ggplot2)
ggplot(myData, aes(pclass, avg_fare)) + geom_bar(stat="identity")
```



## More challenging

11. Create a data frame with:
- `pclass`
- `sex`
- `age`
- `fare`
- `survived`
- `parch`
Only for people who did not survive. 

```{r}
query <- "select pclass, sex, age, fare, survived, parch from titanic_king where survived=0;"

died <- dbGetQuery(con, query)
```

12. Create a graph of the average survival over the different classes. Make two distinct lines for male and female passengers. Create this plot using ggplot. 

```{r}
query <- 'select sex, pclass, avg(survived) as percent_survived from titanic_king
group by sex, pclass;'

survival <- dbGetQuery(con, query)

ggplot(survival, aes(x = pclass, y = percent_survived, group = sex)) + geom_line(aes(colour = sex))
```
