---
title: "Project 4"
author: "Menghan Hu"
date: "Due October 7, 2016 at 11:59pm"
output:
  html_document


---

<style type="text/css">
.table {

    width: 80%;
    margin-left:10%; 
    margin-right:10%;
}
</style>
```{r,setup, echo=FALSE, cache=TRUE}
## numbers >= 10^5 will be denoted in scientific notation,
## and rounded to 2 digits
options(scipen = 3, digits = 3)
```




#Project Goals:

The goal of this project is to get aquainted with the Rstudio server as well as to work on creating functions. You will do some function refactoring as well.[Lab 4 - Markdown]()


# Part I: Function Building with the jackknife function


## Jackknife of the Mean!

- Have an estimator $\hat{\theta}$ of parameter $\theta$  
  want the standard error of our estimate, $se_{\hat{\theta}}$
- The jackknife approximation:
    + omit case $i$, get estimate $\hat{\theta}_{(-i)}$
    + Take the variance of all the $\hat{\theta}_{(-i)}$
    + multiply that variance by $\frac{(n-1)^2}{n}$ to get $\approx$ variance of $\hat{\theta}$
- then $se_{\hat{\theta}}=$ square root of that variance

<small>(Why $(n-1)^2 / n$?  Think about just getting the standard
    error of the mean)</small>



Write a function called `mean.jackknife` that takes argument `a_vector` and returns a jackknife estimate of the standard error.


```{r, eval = FALSE}
mean.jackknife <- function(a_vector) {
  n <- length(a_vector)
  mean.list <- c()
  # Put your function in here
  for (i in 1:n){
    
    mean.list <- c(mean.list, mean(a_vector[-i]))
    
  }
  jackknife.stderr <- sqrt((n-1)^2/n * var(mean.list))
  return(jackknife.stderr)
}
```



#### Jackknife for the mean Example: test your code here


```{r, eval = FALSE}
set.seed(100)
some_normals <- rnorm(100,mean=7,sd=5)
mean(some_normals)
(formula_se_of_mean <- sd(some_normals)/sqrt(length(some_normals)))
all.equal(formula_se_of_mean,mean.jackknife(some_normals))
```



##Jackknife for Gamma Parameters


Recall our friend the method of moments estimator:

```{r, eval = FALSE}
gamma.est <- function(the_data) {
  m <- mean(the_data)
  v <- var(the_data)
  a <- m^2/v
  s <- v/m
  return(c(a=a,s=s))
}
```


###Jackknife for Gamma Parameters Function

Write a function called `gamma.jackknife` that takes argument `a_vector` and returns jackknife standard error estimates on the gamma parameters. 

```{r, eval = FALSE}

gamma.jackknife <- function(a_vector) {
  n <- length(a_vector)
  
  a.list <- c()
  s.list <- c()

  for (i in 1:n){
    est <- gamma.est(a_vector[-i])
    a.list <- c(a.list,est[1])
    s.list <- c(s.list,est[2])
  }
  
  jackknife.stderrs <- c(a.jkstd = sqrt((n-1)^2/n * var(a.list)), 
                         s.jkstd = sqrt((n-1)^2/n * var(s.list)))
  
  return(jackknife.stderrs)
}
```



###Jackknife for Gamma Parameters Example


```{r, eval = FALSE}
set.seed(100)
input <- rgamma(1000, shape=0.1, scale=10)
gamma.est(input)
gamma.jackknife(input)
```



##Jackknife for linear regression coefficients

Write a function called `jackknife.lm` that takes arguments `df`, `formula` and `p` and returns jackknife standard error estimates on the coefficients of a linear regression model. 

```{r, eval = FALSE}
jackknife.lm <- function(df,formula,p) {
  
  n <- nrow(df)
  jackknife.est <- matrix(0, nrow = p, ncol = n)
  
  for (i in 1:n){
    my.lm <- lm(formula = formula, data = df[-i, ])
    
    jackknife.est[,i] <- summary(my.lm)$coef[,1]
    
  }
  
  jackknife.stderr <- apply(jackknife.est, 1, 
                            function(x) sqrt((n-1)^2/n * var(x)))
  
  
  return(jackknife.stderr)
}
```



###Jackknife for linear regression coefficients Example


```{r, eval=FALSE}
output <- 1.2 + 0.6*input +  rnorm(1000, 0, 2.1)
data <- data.frame(output,input)
my.lm <- lm(output~input, data=data)
coefficients(my.lm)
# "Official" standard errors
sqrt(diag(vcov(my.lm)))
jackknife.lm(df=data,formula="output~input",p=2)
```




##Refactoring the Jackknife


- Omitting one point or row is a common sub-task

- The general pattern:

```
figure out the size of the data
for each case
   omit that case
   repeat some estimation and get a vector of numbers
take variances across cases
scale up variances
take the square roots
```

- Refactor by extracting the common "omit one" operation

- Refactor by defining a general "jackknife" operation



###The Common Operation


- _Problem_: Omit one particular data point from a larger structure

- _Difficulty_: Do we need a comma in the index or not?

- _Solution_: Works for vectors, lists, 1D and 2D arrays, matrices, data frames:



## Goal: 


- Make the function select the correct dimensions
    - length for a 1d object
    - number of rows for 2d
- Write a function `omit.case` that omits a point given the data and returns the data minus that point. Make sure it can handle higher dimensions. 
    ```{r, eval = FALSE}
          omit.case <- function(the_data,omitted_point) {
        # This should take the data and omit one point at a time and return the new data
            
            
            if(is.null(dim(the_data))){
              jk.data <- the_data[-omitted_point]
            } else{
              jk.data <- the_data[-omitted_point,]
            }
            
            return(jk.data)
            }
    ```
- Write a function `omit_and_est` that takes the data with an omitted point and returns whatever function your estimator does.  
    ```{r, eval = FALSE}
    
        omit_and_est <- function(omit) {
              # This function should take the output of omit.case and use it as input for the estimator
          estimator(omit.case(the_data, omit))
        }
    ```


```{r, eval = FALSE}
jackknife <- function(estimator,the_data) {
  
  # Define n to be lenth or number of rows here
  if (is.null(dim(the_data))){
    n <- length(the_data)
  } else {
    n <- nrow(the_data)
  }
  
  omit_and_est <- function(omit) {
              # This function should take the output of omit.case and use it as input for the estimator
          estimator(omit.case(the_data, omit))
        }
  
  
  jackknife.ests <- matrix(sapply(1:n, omit_and_est), ncol=n)
  var.of.reestimates <- apply(jackknife.ests,1,var)
  jackknife.var <- ((n-1)^2/n)* var.of.reestimates
  jackknife.stderr <- sqrt(jackknife.var)
  return(jackknife.stderr)
}
```



---

###It works


```{r, eval = FALSE}
jackknife(estimator=mean,the_data=some_normals)
all.equal(jackknife(estimator=mean,the_data=some_normals),
          mean.jackknife(some_normals))
```

---

```{r, eval = FALSE}
all.equal(jackknife(estimator=gamma.est,the_data=data$input),
          gamma.jackknife(data$input))
```

---


```{r, eval = FALSE}
all.equal(jackknife(estimator=gamma.est,the_data=data$input),
          gamma.jackknife(data$input), check.names=FALSE)
```




```{r, eval = FALSE}
est.coefs <- function(the_data) {
  return(lm(output~input,data=the_data)$coefficients)
}
est.coefs(data)
all.equal(est.coefs(data), coefficients(my.lm))
```

---


```{r, eval=FALSE}
jackknife(estimator=est.coefs,the_data=data)
all.equal(jackknife(estimator=est.coefs,the_data=data),
          jackknife.lm(df=data,formula="output~input",p=2))
```


##Further Refactoring of jackknife()


The code for `jackknife()` is still a bit clunky:
- Ugly `if-else` for finding `n`
- Bit at the end for scaling variances down to standard errors

- write a function that calculates the `n` needed for the above code:


    ```{r, eval=FALSE}
      data_size <- function(the_data) {
      
        if (is.null(dim(the_data))){
    n <- length(the_data)
         } else {
    n <- nrow(the_data)
      }
  return(n)
      }
    ```
- Write a funcrion that calculate the variance of all the estimates and returns the standard error
```{r, eval=FALSE}
scale_and_sqrt_vars <- function(jackknife.ests,n) {
  jackknife.stderr <- apply(jackknife.est, 1, 
                            function(x) sqrt((n-1)^2/n * var(x)))
  return(jackknife.stderr)
}
```



###Now invoke those functions

```{r, eval=FALSE}
jackknife <- function(estimator,the_data) {
  n <- data_size(the_data)
  omit_and_est <- function(omit) {
    estimator(omit.case(the_data,omit))
  }
  jackknife.ests <- matrix(sapply(1:n, omit_and_est), ncol=n)
  return(scale_and_sqrt_vars(jackknife.ests,n))
}
```

